---
title: "Coding for Final Project"
output: html_document
date: "2025-04-17"
---
```{r}
library(questionr)
library(tidyr)
library(dplyr)
library(ggplot2)
library(dunn.test)
```

Descriptive statistics. 
Count tables for nominal variables Pregnancy Diabetes, High Blood Pressure, Race. This creates counts of each variable That I can use as a general description of my variables.
```{r}
count(Cesarean_FinalProject, Race)
count(Cesarean_FinalProject, High_Blood_Pressure)
count(Cesarean_FinalProject, Pregnancy_Diabetes)
```
Frequency table for ordinal variables BMIcat, and Agecat. This produces counts and percentages of the ordinal variables. 
```{r}
freq(Cesarean_FinalProject$BMIcat, cum=TRUE)
freq(Cesarean_FinalProject$Agecat, cum=TRUE)
```
Descriptive statistics for continuous variable Cesarian deliveries and vaginal deliveries that gives me the mean, standard deviation, and count of each. 
```{r}
Cesarean_FinalProject%>% 
  drop_na(Cesarean_Birth)%>%
  summarize(mean(Cesarean_Birth), sd(Cesarean_Birth), count=n())

Cesarean_FinalProject%>% 
  drop_na(Vaginal_Birth)%>%
  summarize(mean(Vaginal_Birth), sd(Vaginal_Birth), count=n())

```

Descriptive statistics of continuous variables by nominal or ordinal groups to find mean, standard deviation, and count on race with cesarean and vaginal births 
```{r}
Cesarean_FinalProject%>%
  group_by(Race)%>%
  drop_na(Cesarean_Birth)%>%
  summarize(mean(Cesarean_Birth), sd(Cesarean_Birth), count=n())

Cesarean_FinalProject%>%
  group_by(Race)%>%
  drop_na(Vaginal_Birth)%>%
  summarize(mean(Vaginal_Birth), sd(Vaginal_Birth), count=n())
```
Descriptive statistics of continuous variables by nominal or ordinal groups to find mean, standard deviation, and count on high blood pressure with cesarean birth 
```{r}
Cesarean_FinalProject%>%
  group_by(High_Blood_Pressure)%>%
  drop_na(Cesarean_Birth)%>%
  summarize(mean(Cesarean_Birth), sd(Cesarean_Birth), count=n())
```
Descriptive statistics of continuous variables by nominal or ordinal groups to find mean, standard deviation, and count on diabetes during pregnancy with cesarean birth 
```{r}
Cesarean_FinalProject%>%
  group_by(Pregnancy_Diabetes)%>%
  drop_na(Cesarean_Birth)%>%
  summarize(mean(Cesarean_Birth), sd(Cesarean_Birth), count=n())
```

Descriptive statistics to find mean, standard deviation, and count from BMIcat in cesarean birth. 
```{r}
Cesarean_FinalProject%>%
  group_by(BMIcat)%>%
  drop_na(Cesarean_Birth)%>%
  summarize(mean(Cesarean_Birth), sd(Cesarean_Birth), count=n())
```
Descriptive statistics to find mean, standard deviation, and count from agecat in cesarean birth. The caveat with age is that there might be a discrepancy with the age in which people took part in the NHANES data and when people actually had a cesarean/ vaginal birth. Potentially why there are so many 45+ people in this data set is because the age for the data set if way considered off. 
```{r}
Cesarean_FinalProject%>%
  group_by(Agecat)%>%
  drop_na(Cesarean_Birth)%>%
  summarize(mean(Cesarean_Birth), sd(Cesarean_Birth), count=n())
```
```{r}
#this is only for the t-test
Cesarean_FinalProject_reduced <- Cesarean_FinalProject %>% 
  filter(Pregnancy_Diabetes == "No" | Pregnancy_Diabetes == "Yes") %>% 
  filter(High_Blood_Pressure == "No" | High_Blood_Pressure == "Yes")
Cesarean_FinalProject_reduced
```
T-Test for Pregnancy diabetes and BMI and High blood pressure and BMI
```{r}
hist(Cesarean_FinalProject_reduced$BMI)
Cesarean_FinalProject_reduced%>%
  drop_na(BMI)%>%
  group_by(Pregnancy_Diabetes)%>%
  summarize(mean(BMI), sd(BMI), count=n())

#var.test(Contvar~Catvar, data=, alternative="two.sided")
var.test(BMI~Pregnancy_Diabetes, data=Cesarean_FinalProject_reduced, alternative="two.sided")
t.test(BMI~Pregnancy_Diabetes, var.equal=FALSE, data=Cesarean_FinalProject_reduced)
```
There is a statistical difference with Pregnancy Diabetes and BMI with a p-value of 0.02124

T-test with High blood pressure with BMI 
```{r}
hist(Cesarean_FinalProject_reduced$BMI)
Cesarean_FinalProject_reduced%>%
  drop_na(BMI)%>%
  group_by(High_Blood_Pressure)%>%
  summarize(mean(BMI), sd(BMI), count=n())

#var.test(Contvar~Catvar, data=, alternative="two.sided")
var.test(BMI~High_Blood_Pressure, data=Cesarean_FinalProject_reduced, alternative="two.sided")
t.test(BMI~High_Blood_Pressure,var.equal=TRUE, data=Cesarean_FinalProject_reduced)
```
High blood pressure and BMI is statistically significant as the p-value is 1.728e-11. 

Kruskal-Wallis Dunn test Continuous variable w/ three or more categorical variables)
Cesarean Birth and race 
```{r}

#1. Generate summary statistics of cesarean birth and race?
Cesarean_FinalProject%>%
  drop_na(Cesarean_Birth)%>%
  group_by(Race)%>%
  summarize(mean(Cesarean_Birth), sd (Cesarean_Birth), count=n())
  
#2. Generate a boxplot of cesarean birth and race to visualize the summary statistics
ggplot(Cesarean_FinalProject, aes(x=Race, y=Cesarean_Birth)) + geom_boxplot(fill= "darkgreen") + 
  labs(x="Race", y="Cesarean Birth", title="Race/Cesarean Birth")
#3. Run the Kruskal-Wallis test and the post-hoc Dunn test to see if there are any significant differences in cesarean birth race race. The code for the Kruskal-Wallis and Dunn test is: 
#dunn.test(x=data$contvar, g=data$catvar, kw=TRUE)
dunn.test(x=Cesarean_FinalProject$Cesarean_Birth, g=Cesarean_FinalProject$Race, kw=TRUE)

```
Cesarean birth and High blood pressure
```{r}
Cesarean_FinalProject%>%
  drop_na(Cesarean_Birth)%>%
  group_by(High_Blood_Pressure)%>%
  summarize(mean(Cesarean_Birth), sd (Cesarean_Birth), count=n())
  
#2. Generate a boxplot of cesarean birth and race to visualize the summary statistics
ggplot(Cesarean_FinalProject, aes(x=High_Blood_Pressure, y=Cesarean_Birth)) + geom_boxplot(fill= "lightblue") + 
  labs(x="High Blood Pressure", y="Cesarean Birth", title="High Blood Pressure/Cesarean Birth")
#3. Run the Kruskal-Wallis test and the post-hoc Dunn test to see if there are any significant differences in cesarean birth race race. The code for the Kruskal-Wallis and Dunn test is: 
#dunn.test(x=data$contvar, g=data$catvar, kw=TRUE)
dunn.test(x=Cesarean_FinalProject$Cesarean_Birth, g=Cesarean_FinalProject$High_Blood_Pressure, kw=TRUE)
```

Cesarean birth and pregnancy diabetes 
```{r}
Cesarean_FinalProject%>%
  drop_na(Cesarean_Birth)%>%
  group_by(Pregnancy_Diabetes)%>%
  summarize(mean(Cesarean_Birth), sd (Cesarean_Birth), count=n())
  
#2. Generate a boxplot of cesarean birth and race to visualize the summary statistics
ggplot(Cesarean_FinalProject, aes(x=Pregnancy_Diabetes, y=Cesarean_Birth)) + geom_boxplot(fill= "blue") + 
  labs(x="Pregnancy Diabetes", y="Cesarean Birth", title="Pregnancy Diabetes/Cesarean Birth")
#3. Run the Kruskal-Wallis test and the post-hoc Dunn test to see if there are any significant differences in cesarean birth race race. The code for the Kruskal-Wallis and Dunn test is: 
#dunn.test(x=data$contvar, g=data$catvar, kw=TRUE)
dunn.test(x=Cesarean_FinalProject$Cesarean_Birth, g=Cesarean_FinalProject$Pregnancy_Diabetes, kw=TRUE)
```
Chi Square 
high blood pressure and race
```{r}
#a. Generate a two-way frequency table of pregnacy diabetes and :
#table_name <- xtabs(~ CatVar1 + CatVar2, data=)
#print the table
#show the table in percentages: prop.table(table_name)*100
HBP_Race_Table <- xtabs(~ High_Blood_Pressure + Race, data=Cesarean_FinalProject)
HBP_Race_Table
prop.table(HBP_Race_Table)*100
#b. run the chi-square test:
#chisq.test(table_name)
chisq.test(HBP_Race_Table)
```
Pregnancy diabetes, BMIcat
```{r}
PB_BMIcat_Table <- xtabs(~ Pregnancy_Diabetes + BMIcat, data=Cesarean_FinalProject)
PB_BMIcat_Table
prop.table(PB_BMIcat_Table)*100
chisq.test(PB_BMIcat_Table)
```
Correlation Statitics 
```{r}
#scatter plot code: #ggplot(aes(x=DepVar, y=IndVar), data=) + geom_point(color="") + geom_smooth(method=lm, color="") + labs(x="", y="") + theme_bw()
Cesarean_FinalProject %>% 
  ggplot(.,aes(Cesarean_Birth,Vaginal_Birth)) +
  geom_point(color="lightblue") +
  geom_smooth(method=lm, color="navy") +
  theme_bw()
#correlation test code: cor.test(data$DepVar, data$IndVar, method="pearson")
cor.test(Cesarean_FinalProject$Cesarean_Birth,Cesarean_FinalProject$Vaginal_Birth)
```
Linear Regression models, Cesarean Birth and BMI & Vaginal Birth and BMI 
```{r}
#model_name <- lm(DepVar~IndVar, data=)
#summary(model_name)
Cesarean_BMI_Regression <- lm(BMI~Cesarean_Birth, data= Cesarean_FinalProject)
summary(Cesarean_BMI_Regression)

Vaginal_BMI_Regression <- lm(BMI~Vaginal_Birth, data= Cesarean_FinalProject)
summary(Vaginal_BMI_Regression)
```
Create yes no for cesarean birth for odds ratio 
```{r}
Cesarean_FinalProject_reduced$Cesarean_BirthYN[Cesarean_FinalProject_reduced$Cesarean_Birth==1] <- "Yes"
Cesarean_FinalProject_reduced$Cesarean_BirthYN[Cesarean_FinalProject_reduced$Cesarean_Birth==2] <- "Yes"
Cesarean_FinalProject_reduced$Cesarean_BirthYN[Cesarean_FinalProject_reduced$Cesarean_Birth==3] <- "Yes"
Cesarean_FinalProject_reduced$Cesarean_BirthYN[Cesarean_FinalProject_reduced$Cesarean_Birth==4] <- "Yes"
Cesarean_FinalProject_reduced$Cesarean_BirthYN[Cesarean_FinalProject_reduced$Cesarean_Birth==5] <- "Yes"
Cesarean_FinalProject_reduced$Cesarean_BirthYN[Cesarean_FinalProject_reduced$Cesarean_Birth==6] <- "Yes"
Cesarean_FinalProject_reduced$Cesarean_BirthYN[Cesarean_FinalProject_reduced$Cesarean_Birth==7] <- "Yes"
Cesarean_FinalProject_reduced$Cesarean_BirthYN[Cesarean_FinalProject_reduced$Cesarean_Birth==0] <- "No"
factor(c(Cesarean_FinalProject_reduced$Cesarean_BirthYN, levels=c("No", "Yes")))
head(Cesarean_FinalProject_reduced)
```
Create odds ratio with Cesarean births and race
```{r}
#1. Create the odds table:
#odds_table_name <- table(data$IndVar, data$DepVar)
Race_Cesarean_odds_table <- table(Cesarean_FinalProject_reduced$Race, Cesarean_FinalProject_reduced$Cesarean_BirthYN)
#2. Run the oddsratio on the table you created:
#oddsratio(odds_table_name)
oddsratio(Race_Cesarean_odds_table)
#3. Graph the results in a mosaic table:
#ggplot(data=) + geom_mosaic(aes(x=product(IndVar), fill=DepVar), na.rm=TRUE, show.legend=FALSE) + labs(title="", x="", y="")
ggplot(data=Cesarean_FinalProject_reduced)+ 
  geom_mosaic(aes(x=product(Race), fill=Cesarean_BirthYN), 
              na.rm=TRUE, show.legend=FALSE) +
  labs (title= "Cesarean By Race", x="Race", y="Yes/No")
```

odds ratio with cesarean birth and age catagoriy 
```{r}
#odds_table_name <- table(data$IndVar, data$DepVar)
Agecat_Cesarean_odds_table <- table(Cesarean_FinalProject_reduced$Agecat, Cesarean_FinalProject_reduced$Cesarean_BirthYN)
#2. Run the oddsratio on the table you created:
#oddsratio(odds_table_name)
oddsratio(Agecat_Cesarean_odds_table)
#3. Graph the results in a mosaic table:
#ggplot(data=) + geom_mosaic(aes(x=product(IndVar), fill=DepVar), na.rm=TRUE, show.legend=FALSE) + labs(title="", x="", y="")
ggplot(data=Cesarean_FinalProject_reduced)+ 
  geom_mosaic(aes(x=product(Agecat), fill=Cesarean_BirthYN), 
              na.rm=TRUE, show.legend=FALSE) +
  labs (title= "Cesarean By Age cat", x="Race", y="Yes/No")
```

```{r}
#1. Create the odds table:
#odds_table_name <- table(data$IndVar, data$DepVar)
HBP_Cesarean_odds_table <- table(Cesarean_FinalProject_reduced$BMIcat, Cesarean_FinalProject_reduced$Cesarean_BirthYN)
#2. Run the oddsratio on the table you created:
#oddsratio(odds_table_name)
oddsratio(HBP_Cesarean_odds_table)
#3. Graph the results in a mosaic table:
#ggplot(data=) + geom_mosaic(aes(x=product(IndVar), fill=DepVar), na.rm=TRUE, show.legend=FALSE) + labs(title="", x="", y="")
ggplot(data=Cesarean_FinalProject_reduced)+ 
  geom_mosaic(aes(x=product(BMIcat), fill=Cesarean_BirthYN), 
              na.rm=TRUE, show.legend=FALSE) +
  labs (title= "Cesarean By BMIcat", x="BMIcat", y="Yes/No")
```


Yes NO for Vaginal birth for compairative odds ratio. 
```{r}
Cesarean_FinalProject_reduced$Vaginal_BirthYN[Cesarean_FinalProject_reduced$Vaginal_Birth==1] <- "Yes"
Cesarean_FinalProject_reduced$Vaginal_BirthYN[Cesarean_FinalProject_reduced$Vaginal_Birth==2] <- "Yes"
Cesarean_FinalProject_reduced$Vaginal_BirthYN[Cesarean_FinalProject_reduced$Vaginal_Birth==3] <- "Yes"
Cesarean_FinalProject_reduced$Vaginal_BirthYN[Cesarean_FinalProject_reduced$Vaginal_Birth==4] <- "Yes"
Cesarean_FinalProject_reduced$Vaginal_BirthYN[Cesarean_FinalProject_reduced$Vaginal_Birth==5] <- "Yes"
Cesarean_FinalProject_reduced$Vaginal_BirthYN[Cesarean_FinalProject_reduced$Vaginal_Birth==6] <- "Yes"
Cesarean_FinalProject_reduced$Vaginal_BirthYN[Cesarean_FinalProject_reduced$Vaginal_Birth==7] <- "Yes"
Cesarean_FinalProject_reduced$Vaginal_BirthYN[Cesarean_FinalProject_reduced$Vaginal_Birth==8] <- "Yes"
Cesarean_FinalProject_reduced$Vaginal_BirthYN[Cesarean_FinalProject_reduced$Vaginal_Birth==9] <- "Yes"
Cesarean_FinalProject_reduced$Vaginal_BirthYN[Cesarean_FinalProject_reduced$Vaginal_Birth==10] <- "Yes"
Cesarean_FinalProject_reduced$Vaginal_BirthYN[Cesarean_FinalProject_reduced$Vaginal_Birth==11] <- "Yes"
Cesarean_FinalProject_reduced$Vaginal_BirthYN[Cesarean_FinalProject_reduced$Vaginal_Birth==0] <- "No"
factor(c(Cesarean_FinalProject_reduced$Vaginal_BirthYN, levels=c("No", "Yes")))
head(Cesarean_FinalProject_reduced)
```
odds ratio for vaginal birth and race 
```{r}
#1. Create the odds table:
#odds_table_name <- table(data$IndVar, data$DepVar)
Race_Vaginal_odds_table <- table(Cesarean_FinalProject_reduced$Race, Cesarean_FinalProject_reduced$Vaginal_BirthYN)
#2. Run the oddsratio on the table you created:
#oddsratio(odds_table_name)
oddsratio(Race_Vaginal_odds_table)
#3. Graph the results in a mosaic table:
#ggplot(data=) + geom_mosaic(aes(x=product(IndVar), fill=DepVar), na.rm=TRUE, show.legend=FALSE) + labs(title="", x="", y="")
ggplot(data=Cesarean_FinalProject_reduced)+ 
  geom_mosaic(aes(x=product(Race), fill=Vaginal_BirthYN), 
              na.rm=TRUE, show.legend=FALSE) +
  labs (title= "Vaginal By Race", x="Race", y="Yes/No")
```
Odds ratio for Vaginal birth and age catagories 
```{r}
#1. Create the odds table:
#odds_table_name <- table(data$IndVar, data$DepVar)
Race_Vaginal_odds_table <- table(Cesarean_FinalProject_reduced$Agecat, Cesarean_FinalProject_reduced$Vaginal_BirthYN)
#2. Run the oddsratio on the table you created:
#oddsratio(odds_table_name)
oddsratio(Race_Vaginal_odds_table)
#3. Graph the results in a mosaic table:
#ggplot(data=) + geom_mosaic(aes(x=product(IndVar), fill=DepVar), na.rm=TRUE, show.legend=FALSE) + labs(title="", x="", y="")
ggplot(data=Cesarean_FinalProject_reduced)+ 
  geom_mosaic(aes(x=product(Agecat), fill=Vaginal_BirthYN), 
              na.rm=TRUE, show.legend=FALSE) +
  labs (title= "Vaginal By Agecat", x="Race", y="Yes/No")
```

Basic graphs for descriptive data. 
```{r}
ggplot(Cesarean_FinalProject, aes(x=Cesarean_Birth)) + geom_histogram(color="lightblue", fill= "steelblue") + 
  labs(title="Cesarean Birth", x="Cesarean Birth", y="Count")
```


```{r}
ggplot(Cesarean_FinalProject, aes(x=Vaginal_Birth)) + geom_histogram(color="lightblue", fill= "steelblue") + 
  labs(title="Vaginal Birth", x="Vaginal Birth", y="Count")
```



